---
output:
  revealjs_presentation:
    self_contained: no
---

# Extended truncated Inverse Gaussian–Poisson model
## (with R examples)

<br>

### Lluís Ramon
### May 9, 2014

  
---

# Summary

<br>

## Parameter extension of tPIG

## Fitting a case example


# Word frequency word data

<br>

<section>

* Modeling word frequency count data
* Studying author's style through its vocabulary

<br>

## tm package (Text Mining)

<br>

```{r, eval = FALSE}
library("tm")
?PlainTextDocument
?removePunctuation
?termFreq
```

---

```{r Paraules capitol 5}
chap5 <- readLines("data/Tirant lo Blanc Chapter 5", encoding = "UTF-8")

library("tm")
chap5 <- PlainTextDocument(chap5)
chap5 <- gsub("'", " ", chap5)
chap5 <- removePunctuation(chap5)

freq_chap5 <- termFreq(chap5)
table(freq_chap5)
```

<br>

* Different versions
* Remove punctuation, capital letters, etc

</section>

# Distributions

<br>

* Inverse Gaussian (IG)
* IG-Poisson mixture
* Truncated IG-Poisson mixture. 
* Extended Truncated IG-Poisson mixture

<br>

## gamlss package

<br>

```{r gamlss, eval = FALSE, message = FALSE}
library(gamlss)
?IG # Inverse Gaussian distribution
?dIG

?PIG # Poisson-inverse Gaussian distribution
?dPIG
```

# Inverse Gaussian

<br>

<section>

$f(x|\mu,\lambda) = \left(\frac{\lambda}{2\pi x^3}\right)^{1/2} \exp\left(\frac{-\lambda (x-\mu)^2}{2 \mu^2 x}\right),  x \in (0,  \infty)$

<br>

* Continuous distribution
* Not a true _"inverse"_

---

```{r IG plot, message = FALSE, echo = FALSE}
library("gamlss")
x <- seq(0.1, 2.5, by = 0.02)
IG1 <- data.frame(x = x, y = dIG(x, 1, 5), dist = "IG(1, 5)", stringsAsFactors= FALSE)
IG2 <- data.frame(x = x, y = dIG(x, 1, 1), dist = "IG(1, 1)", stringsAsFactors= FALSE)
IG3 <- data.frame(x = x, y = dIG(x, 1, 0.1), dist = "IG(1, 0.1)", stringsAsFactors= FALSE)
IG4 <- data.frame(x = x, y = dIG(x, 1, 0.3), dist = "IG(1, 0.3)", stringsAsFactors= FALSE)

IG_prior <- rbind(IG1, IG2, IG3, IG4)

library(ggplot2)
qplot(x, y, data = IG_prior, geom = "line", col = dist)
```

---

```{r, message = FALSE, eval = FALSE}
library("gamlss")
x <- seq(0.1, 2.5, by = 0.02)
IG1 <- data.frame(x = x, y = dIG(x, 1, 5), dist = "IG(1, 5)", stringsAsFactors= FALSE)
IG2 <- data.frame(x = x, y = dIG(x, 1, 1), dist = "IG(1, 1)", stringsAsFactors= FALSE)
IG3 <- data.frame(x = x, y = dIG(x, 1, 0.1), dist = "IG(1, 0.1)", stringsAsFactors= FALSE)
IG4 <- data.frame(x = x, y = dIG(x, 1, 0.3), dist = "IG(1, 0.3)", stringsAsFactors= FALSE)

IG_prior <- rbind(IG1, IG2, IG3, IG4)

library(ggplot2)
qplot(x, y, data = IG_prior, geom = "line", col = dist)
```

</section>

# Inverse Gaussian–Poisson model

<br>

<section>

$e^{\alpha \sqrt{(1 -\theta)}} \sqrt{ \frac{2 \alpha}{\pi}} \frac{(\frac{1}{2} \alpha \theta)^r}{r!} K_{r - \frac{1}{2}}(\alpha)$

<br>

* $K_y$ third kind modified Bessel function
* Mixture of a discrete distribution with a continuous distribution 

---

```{r,  echo = FALSE}
x <- 0:10
PIG1 <- data.frame(x = x, y = dPIG(x, 5, 5), dist = "PIG(5, 5)", stringsAsFactors= FALSE)
PIG2 <- data.frame(x = x, y = dPIG(x, 1, 1), dist = "PIG(1, 1)", stringsAsFactors= FALSE)
PIG3 <- data.frame(x = x, y = dPIG(x, 1, 0.1), dist = "PIG(1, 0.1)", stringsAsFactors= FALSE)
PIG4 <- data.frame(x = x, y = dPIG(x, 1, 0.3), dist = "PIG(1, 0.3)", stringsAsFactors= FALSE)

PIG_prior <- rbind(PIG1, PIG2, PIG3, PIG4)

library(ggplot2)
qplot(x, y, data = PIG_prior, geom = "point", col = dist) + facet_grid(.~ dist)
```

---

```{r, eval = FALSE}
x <- 0:10
PIG1 <- data.frame(x = x, y = dPIG(x, 5, 5), dist = "PIG(5, 5)", stringsAsFactors= FALSE)
PIG2 <- data.frame(x = x, y = dPIG(x, 1, 1), dist = "PIG(1, 1)", stringsAsFactors= FALSE)
PIG3 <- data.frame(x = x, y = dPIG(x, 1, 0.1), dist = "PIG(1, 0.1)", stringsAsFactors= FALSE)
PIG4 <- data.frame(x = x, y = dPIG(x, 1, 0.3), dist = "PIG(1, 0.3)", stringsAsFactors= FALSE)

PIG_prior <- rbind(PIG1, PIG2, PIG3, PIG4)

library(ggplot2)
qplot(x, y, data = PIG_prior, geom = "point", col = dist) + facet_grid(.~ dist)
```

</section>

# truncated Inverse Gaussian–Poisson model

<br>

<section>

* No words written 0 times
* A left truncation in zero needed
* Conditional probability

<br>

$$P(X = k | X > 0) = \frac{P(X = k  \cap  X> 0)}{1 - P(X = 0)}$$

<br>

## gamlss.tr package

```{r, eval = FALSE}
library("gamlss.tr")
trun.d(par = 0, family = "PIG", type = "left")
?trun.p; ?trun.q, ...
```


---

```{r, echo = FALSE}
library("gamlss.tr")
dtPIG <- trun.d(par = 0, family = "PIG", type = "left")

x <- 1:8
tPIG1 <- data.frame(x = x, y = dtPIG(x, 1.5, 1), dist = "tPIG(1.5, 1)", stringsAsFactors= FALSE)
tPIG2 <- data.frame(x = x, y = dtPIG(x, mu = 1.5, sigma = 4), dist = "tPIG(1.5, 2)", stringsAsFactors= FALSE)
tPIG3 <- data.frame(x = x, y = dtPIG(x, 1.5, 0.1), dist = "tPIG(1.5, 0.1)", stringsAsFactors= FALSE)
tPIG4 <- data.frame(x = x, y = dtPIG(x, 1.5, 0.3), dist = "tPIG(1.5, 0.3)", stringsAsFactors= FALSE)

tPIG_prior <- rbind(tPIG1, tPIG2, tPIG3, tPIG4)

library(ggplot2)
qplot(x, y, data = tPIG_prior, geom = "point", col = dist) + facet_grid(.~ dist)
```

---

```{r, eval = FALSE}
library("gamlss.tr")
dtPIG <- trun.d(par = 0, family = "PIG", type = "left")

x <- 1:8
tPIG1 <- data.frame(x = x, y = dtPIG(x, 1.5, 1), dist = "tPIG(1.5, 1)", stringsAsFactors= FALSE)
tPIG2 <- data.frame(x = x, y = dtPIG(x, mu = 1.5, sigma = 4), dist = "tPIG(1.5, 2)", stringsAsFactors= FALSE)
tPIG3 <- data.frame(x = x, y = dtPIG(x, 1.5, 0.1), dist = "tPIG(1.5, 0.1)", stringsAsFactors= FALSE)
tPIG4 <- data.frame(x = x, y = dtPIG(x, 1.5, 0.3), dist = "tPIG(1.5, 0.3)", stringsAsFactors= FALSE)

tPIG_prior <- rbind(tPIG1, tPIG2, tPIG3, tPIG4)

library(ggplot2)
qplot(x, y, data = tPIG_prior, geom = "point", col = dist) + facet_grid(.~ dist)
```

</section>

# Extended truncated Inv..

<section>

<br>

* Extend parametric space $\alpha \in (-1, 0)$
* Extended version = not extended if $\alpha \geq 0$
* Recurrent function for calculating the model

```{r contour plots, fig.height= 4.5, fig.width= 9, echo = FALSE}
p1 <- function(alpha, theta){
  
  0.5*alpha*theta/(exp(alpha*(1 - sqrt(1 -theta))) - 1)
  
}

p2 <- function(alpha, theta){
  
  theta*(1 + alpha)*p1(alpha, theta)/4  
  
}

x <- seq(-1, 1, length.out = 26)
y <- seq(0.75, 1, length.out = 26)

dadesP1 <- sapply(x, FUN = p1, y)
dadesP2 <- sapply(x, FUN = p2, y)

colnames(dadesP1) <- y; row.names(dadesP1) <- x
colnames(dadesP2) <- y; row.names(dadesP2) <- x

par(mfcol = c(1,2))
contour(y, x, dadesP1, main = expression(P[1:n](alpha, theta)))
contour(y, x, dadesP2, main = expression(P[2:n](alpha, theta)))
par(mfcol = c(1,1))
```

---

```{r, fig.height= 4.5, fig.width= 9, eval = FALSE}
p1 <- function(alpha, theta){
  
  0.5*alpha*theta/(exp(alpha*(1 - sqrt(1 -theta))) - 1)
  
}

p2 <- function(alpha, theta){
  
  theta*(1 + alpha)*p1(alpha, theta)/4  
  
}

x <- seq(-1, 1, length.out = 26)
y <- seq(0.75, 1, length.out = 26)

dadesP1 <- sapply(x, FUN = p1, y)
dadesP2 <- sapply(x, FUN = p2, y)

colnames(dadesP1) <- y; row.names(dadesP1) <- x
colnames(dadesP2) <- y; row.names(dadesP2) <- x

par(mfcol = c(1,2))
contour(y, x, dadesP1, main = expression(P[1:n](alpha, theta)))
contour(y, x, dadesP2, main = expression(P[2:n](alpha, theta)))
par(mfcol = c(1,1))
```

</section>

# Personal opinion

<br>

* Full of details article
* Curious about _Not for comercial use_
* Great to have code and data with the article
* Why not a Sichel? generalization with additional parameter
